---
title: "FA 800"
author: "Beatrice Spinola"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(quantmod)
library(lubridate)
library(dplyr)
library(PerformanceAnalytics)
library(MASS)
library(tidyr)
library(class)
library(reshape2)
library(ggplot2)
library(vioplot)
library(sfsmisc)
library(dplyr)
library(corrplot)
library(aod)
```

### Read data from csv
```{r}
model_data <- read.csv('C:/Users/beasp/OneDrive/Desktop/FA 800 - Project in Financial Analytics/etfs_log_rtn.csv')

# Make date column into row label
rownames(model_data) <- model_data[,1]
model_data <- model_data[,-1]
```

### Data Cleaning
```{r}
summary(model_data)

# Convert Post_COVID to a factor, indicating that it should be treated as a categorical variable.
model_data$post_covid <- as.factor(model_data$post_covid)
```

# Exploratory Data Analysis
```{r}
# Summary Stats
summary(model_data)

# Boxplot over entire dataset
boxplot.matrix(as.matrix(model_data[, 1:6]), cex.axis = 0.6)
boxplot.matrix(as.matrix(model_data[, 7:16]), cex.axis = 0.6)

# Reading Sector ETF Data and Formatting
etf_log_rtn <- read.csv("C:/Users/beasp/OneDrive/Desktop/FA 800 - Project in Financial Analytics/etfs_log_rtn.csv")
etf_log_rtn$Date <- as.Date(etf_log_rtn$Date, format="%m/%d/%Y")
class(etf_log_rtn$Date)

# Split data in pre and post-covid
pre_covid_etfs <- filter(etf_log_rtn, Date < '2020-01-10')
post_covid_etfs <- filter(etf_log_rtn, Date >= '2020-01-10')

# Boxplots pre and post
#Pre-covid
boxplot.matrix(as.matrix(filter(model_data, post_covid == 'FALSE')[,1:6]), cex.axis = 0.6)
boxplot.matrix(as.matrix(filter(model_data, post_covid == 'FALSE')[,7:16]), cex.axis = 0.6)

#Post-covid
boxplot.matrix(as.matrix(filter(model_data, post_covid == 'TRUE')[,1:6]), cex.axis = 0.6)
boxplot.matrix(as.matrix(filter(model_data, post_covid == 'TRUE')[,7:16]), cex.axis = 0.6)

# Annualized ETF Log Returns pre and post
mean_rtns <- rbind(colMeans(filter(model_data, post_covid == 'FALSE')[,1:16])*252,
                   colMeans(filter(model_data, post_covid == 'TRUE')[,1:16])*252,
                   colMeans(filter(model_data, post_covid == 'TRUE')[,1:16])*252-colMeans(filter(model_data, post_covid == 'FALSE')[,1:16])*252)
rownames(mean_rtns) <- c('Pre', 'Post', 'Change')
mean_rtns
write.csv(mean_rtns, "mean_rtns.csv")

# Annualized Volatility
vol <- t(cbind(sapply(filter(model_data, post_covid == 'FALSE')[,1:16], sd)*sqrt(252),
               sapply(filter(model_data, post_covid == 'TRUE')[,1:16], sd)*sqrt(252),
               sapply(filter(model_data, post_covid == 'TRUE')[,1:16], sd)*sqrt(252)-sapply(filter(model_data, post_covid == 'FALSE')[,1:16], sd)*sqrt(252)))
rownames(vol) <- c('Pre', 'Post', 'Change')
vol
write.csv(vol, "vol.csv")

# Correlation Matrices
cor(filter(model_data, post_covid == 'FALSE')[,1:6])
cor(filter(model_data, post_covid == 'FALSE')[,7:16])

cor(filter(model_data, post_covid == 'TRUE')[,1:6])
cor(filter(model_data, post_covid == 'TRUE')[,7:16])


par(mfrow = c(1, 2))
corrplot.mixed(cor(filter(model_data, post_covid == 'FALSE')[,1:6]), 
         upper = 'color', is.corr = FALSE, lower.col = COL1('Blues'), upper.col = COL1('Blues'), number.cex = 0.6, tl.cex = 0.6)
corrplot.mixed(cor(filter(model_data, post_covid == 'FALSE')[,7:16]), 
         upper = 'color', is.corr = FALSE, lower.col = COL1('Blues'), upper.col = COL1('Blues'), number.cex = 0.5, tl.cex = 0.6)

corrplot.mixed(cor(filter(model_data, post_covid == 'TRUE')[,1:6]), 
               upper = 'color', is.corr = FALSE, lower.col = COL1('Blues'), upper.col = COL1('Blues'), number.cex = 0.6, tl.cex = 0.6)
corrplot.mixed(cor(filter(model_data, post_covid == 'TRUE')[,7:16]), 
               upper = 'color', is.corr = FALSE, lower.col = COL1('Blues'), upper.col = COL1('Blues'), number.cex = 0.5, tl.cex = 0.6)

?corrplo

# Correlation Matrices
cor_pre_covid_stocks <- cor(filter(model_data, post_covid == 'FALSE')[,1:6])
cor_pre_covid_etfs <- cor(filter(model_data, post_covid == 'FALSE')[,7:16])

cor_post_covid_stocks <- cor(filter(model_data, post_covid == 'TRUE')[,1:6])
cor_post_covid_etfs <- cor(filter(model_data, post_covid == 'TRUE')[,7:16])

cor_pre_covid_stocks
cor_pre_covid_etfs
cor_post_covid_stocks
cor_post_covid_etfs

# This matrix shows if the correlation increased or decreased (-1 is decrease)
cor_diff2 <- (cor_post_covid_stocks - cor_pre_covid_stocks)/abs(cor_post_covid_stocks - cor_pre_covid_stocks)
cor_diff2

cor_diff3 <- (cor_post_covid_etfs - cor_pre_covid_etfs)/abs(cor_post_covid_etfs - cor_pre_covid_etfs)
cor_diff3
```

# Building Classification Models

### Train and Test Datasets
```{r}
set.seed(100)
# Split into train adn test sets
split <- sample(c(rep(0, 0.65 * nrow(model_data)), rep(1, 0.35 * nrow(model_data))))
train_data <- model_data[split == 0, ]
test_data <- model_data[split == 1, ]

# Check how many rows in train and test sets are true or false for target var
# Train Set
nrow(filter(train_data, post_covid == 'FALSE'))
nrow(filter(train_data, post_covid == 'TRUE'))

# Test Set
nrow(filter(test_data, post_covid == 'FALSE'))
nrow(filter(test_data, post_covid == 'TRUE'))
```
\
We are looking at the impact of Covid-19 on six different stock markets. Our plan is to build a model with the sector ETFs as explanatory variables and the binary variable post_covid as the response variable.
\
### KNN
```{r}
# KNN
for (i in c(1, 2, 3, 5, 10, 15, 20)) {
  test.x = as.data.frame(test_data[,c(1)])
  training.x = as.data.frame(train_data[,c(1)])
  knn.pred=knn(training.x, test.x, train_data$post_covid, k=10)
  prediction.knn=cbind(test_data, knn.pred)
  print(i)
  print(table(knn.pred, test_data$post_covid))
  print(mean(knn.pred == test_data$post_covid))
}
```
\
Best value for KNN is K = 3
\
### Final Models
### Logistic Regression
```{r}
# Logistic Regression
# High Change Sectors
glm = glm(post_covid~BVSP+IMOEX.ME+XLE+XLK+XLY, data=train_data, family=binomial)
glm.probs=predict(glm, newdata = test_data, type="response")
glm.pred=rep("FALSE",dim(test_data)[1])
glm.pred[glm.probs>0.5]="TRUE"
table(glm.pred,test_data$post_covid)

# Mid Change Sectors
glm = glm(post_covid~SPY+MXX+XLF+XLB+XLRE+XLU, data=train_data, family=binomial)
glm.probs=predict(glm, newdata = test_data, type="response")
glm.pred=rep("FALSE",dim(test_data)[1])
glm.pred[glm.probs>0.5]="TRUE"
table(glm.pred,test_data$post_covid)

# Low Change Sectors
glm = glm(post_covid~NSEI+IBEX+XLI+XLV+XLP, data=train_data, family=binomial)
glm.probs=predict(glm, newdata = test_data, type="response")
glm.pred=rep("FALSE",dim(test_data)[1])
glm.pred[glm.probs>0.5]="TRUE"
table(glm.pred,test_data$post_covid)
```
\
Logistic Regression could work. Logistic Regression predicted:
High Change: 127/222 rows correctly using the stock index as explanatory variables.
Mid Change: 143/222 rows correctly using the stock index as explanatory variables.
Low Change: 188/222 rows correctly using the stock index as explanatory variables.
\
### LDA
```{r}
# LDA
# High Change Sectors
lda.fit=lda(post_covid~BVSP+IMOEX.ME+XLE+XLK+XLY,data=train_data)
table(predict(lda.fit,test_data)$class, test_data$post_covid)
mean(predict(lda.fit,test_data)$class == test_data$post_covid)

# Mid Change Sectors
lda.fit=lda(post_covid~SPY+MXX+XLF+XLB+XLRE+XLU,data=train_data)
table(predict(lda.fit,test_data)$class, test_data$post_covid)
mean(predict(lda.fit,test_data)$class == test_data$post_covid)

# Low Change Sectors
lda.fit=lda(post_covid~NSEI+IBEX+XLI+XLV+XLP,data=train_data)
table(predict(lda.fit,test_data)$class, test_data$post_covid)
mean(predict(lda.fit,test_data)$class == test_data$post_covid)
```
\
LDA could work. LDA predicted:
High Change: 127/222 rows correctly using the stock index as explanatory variables.
Mid Change: 143/222 rows correctly using the stock index as explanatory variables.
Low Change: 188/222 rows correctly using the stock index as explanatory variables.
\
### QDA
```{r}
# QDA
# High Change Sectors
qda.fit=qda(post_covid~BVSP+IMOEX.ME+XLE+XLK+XLY, data=train_data)
table(predict(qda.fit,test_data)$class, test_data$post_covid)
mean(predict(qda.fit,test_data)$class == test_data$post_covid)

# Mid Change Sectors
qda.fit=qda(post_covid~SPY+MXX+XLF+XLB+XLRE+XLU, data=train_data)
table(predict(qda.fit,test_data)$class, test_data$post_covid)
mean(predict(qda.fit,test_data)$class == test_data$post_covid)

# Low Change Sectors
qda.fit=qda(post_covid~NSEI+IBEX+XLI+XLV+XLP, data=train_data)
table(predict(qda.fit,test_data)$class, test_data$post_covid)
mean(predict(qda.fit,test_data)$class == test_data$post_covid)
```
\
QDA predicted: 
High Change: 107/222 rows correctly using the stock index as explanatory variables.
Mid Change: 101/222 rows correctly using the stock index as explanatory variables.
Low Change: 74/222 rows correctly using the stock index as explanatory variables.
\

\
### KNN
\
```{r}
# KNN
# High Change Sectors
test.x = test_data[,c(3, 4, 7, 10, 15)]
training.x = train_data[,c(3, 4, 7, 10, 15)]
knn.pred=knn(training.x, test.x, train_data$post_covid, k=3)
prediction.knn=cbind(test_data, knn.pred)
table(knn.pred, test_data$post_covid)
mean(knn.pred == test_data$post_covid)

# Mid Change Sector
test.x = test_data[,c(1, 6, 8, 12, 13, 14)]
training.x = train_data[,c(1, 6, 8, 12, 13, 14)]
knn.pred=knn(training.x, test.x, train_data$post_covid, k=3)
prediction.knn=cbind(test_data, knn.pred)
table(knn.pred, test_data$post_covid)
mean(knn.pred == test_data$post_covid)

# Low Change Sectors
test.x = test_data[,c(2, 5, 9, 11, 16)]
training.x = train_data[,c(2, 5, 9, 11, 16)]
knn.pred=knn(training.x, test.x, train_data$post_covid, k=3)
prediction.knn=cbind(test_data, knn.pred)
table(knn.pred, test_data$post_covid)
mean(knn.pred == test_data$post_covid)
```
\
KNN predicted:
High Change: 119/222 rows correctly using the stock index as explanatory variables.
Mid Change: 129/222 rows correctly using the stock index as explanatory variables.
Low Change: 104/222 rows correctly using the stock index as explanatory variables.
\